quollr: An R Package for Visualizing 2D Models from Nonlinear Dimension Reductions in High Dimensional Space

### 2024-03-31 paper-quollr.Rmd
Add about main functions

```{r load-libraries}
library(quollr)
library(tibble)
library(knitr)
library(kableExtra)
library(ggplot2)
library(dplyr)

@@

>>>>

```{r load-libraries}
library(quollr)
library(tibble)
library(knitr)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(ggbeeswarm)

@@

```{r}

## Decide by looking at MSE plot
num_bins_x_s_curve <- 6
num_bins_y_s_curve <- 14
hex_size_s_curve <- 0.13
hb_obj_s_curve <- hex_binning(data = s_curve_noise_umap, x = "UMAP1", 
                      y = "UMAP2", num_bins_x = num_bins_x_s_curve, 
                      num_bins_y = num_bins_y_s_curve, x_start = NA, y_start = NA, 
                      buffer_x = NA, buffer_y = NA, hex_size = hex_size_s_curve, col_start = "UMAP")
## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$centroids))
## Generate all coordinates of hexagons
hex_grid <- as.data.frame(do.call(cbind, hb_obj_s_curve$hex_poly))
## To obtain the standardise counts within hexbins
counts_df <- as.data.frame(do.call(cbind, hb_obj_s_curve$std_cts))
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)
## remove low-density hexagons
hex_grid_with_counts <- dplyr::left_join(hex_grid, counts_df, by = c("hex_poly_id" = "hb_id"))
hex_full_count_df_loop2 <- hex_grid_with_counts |>
  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high"))
hex_full_count_df_filtered <- df_bin_centroids |>
  dplyr::mutate(type = if_else(std_counts <= 0.2, "low", "high")) |>
  dplyr::filter(type == "low")
p2_n <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) #+
  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
p2_n_rm <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  geom_text(data = hex_full_count_df_filtered, color = "red", aes(x = c_x, y = c_y, group = hexID, label = "X",
size = 0.5)) +
  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) #+
  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
# ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
#   geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
#   geom_text(data = all_centroids_df, aes(x = c_x, y = c_y, label = hexID)) +
#   scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
#   coord_fixed()
umap_data_with_hb_id <- as.data.frame(do.call(cbind, hb_obj_s_curve$data_hb_id))

model_object <- fit_highd_model( training_data = s_curve_noise_training, 
                                 nldr_df_with_id = s_curve_noise_umap, 
                                 x = "UMAP1", y = "UMAP2", 
                                 num_bins_x = num_bins_x_s_curve, 
                                 num_bins_y = num_bins_y_s_curve, 
                                 x_start = NA, y_start = NA, 
                                 buffer_x = NA, buffer_y = NA, 
                                 hex_size = hex_size_s_curve,
                                 is_rm_lwd_hex = FALSE, 
                                 benchmark_to_rm_lwd_hex = NA, 
                                 col_start_2d = "UMAP", 
                                 col_start_highd = "x")
df_bin_centroids_s_curve <- model_object$df_bin_centroids
df_bin_s_curve <- model_object$df_bin
cell_count_plot <- ggplot(df_bin_centroids_s_curve, aes(x = reorder(as.factor(hexID), -std_counts), y =
std_counts)) +
  geom_quasirandom() + xlab("hexagonal id") + ylab("Standardized cell count") +
  geom_hline(yintercept = 0.2, colour = "#de2d26") +
  theme(axis.text = element_text(size = 5),
        axis.title = element_text(size = 7),
        axis.text.x = element_text(angle = 90))
benchmark_value_rm_lwd <- stats::quantile(df_bin_centroids_s_curve$std_counts, 
                probs = c(0,0.25,0.5,0.75,1), names = FALSE)[2]
df_bin_centroids_low <- df_bin_centroids_s_curve |>
  dplyr::filter(std_counts <= benchmark_value_rm_lwd)
identify_rm_bins <- find_low_dens_hex(df_bin_centroids_all = df_bin_centroids_s_curve, 
                                      num_bins_x = num_bins_x_s_curve, 
                                      df_bin_centroids_low = df_bin_centroids_low)
df_bin_centroids_s_curve <- df_bin_centroids_s_curve |>
  dplyr::filter(!(hexID %in% identify_rm_bins))
pcentroid_plot_s_curve <- ggplot(data = hex_grid, aes(x = x, y = y)) +
  geom_polygon(fill = "white", color = "black", aes(group = hex_poly_id)) +
  geom_point(data = df_bin_centroids_s_curve, aes(x = c_x, y = c_y), color = "#33a02c") +
  coord_equal() +
  theme_linedraw() +
  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
rm_centroid_df <- df_bin_centroids_low |> 
  dplyr::filter(hexID %in% identify_rm_bins)
p2_n_rm_correct <- ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id, fill = std_counts)) +
  geom_text(data = hex_full_count_df_filtered[6, ], color = "red", aes(x = c_x, y = c_y, group = hexID, label =
"X", size = 0.5)) +
  #geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  #coord_cartesian(xlim =c(-5, 7), ylim = c(-10, 10)) +
  theme_void() +
  theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5,
vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=8), #change legend title font size
        legend.text = element_text(size=6)) +
  guides(fill = guide_colourbar(title = "Standardized count")) #+
  #annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
## Triangulate bin centroids
tr1_object_s_curve <- tri_bin_centroids(df_bin_centroids_s_curve, x = "c_x", y = "c_y")
tr_from_to_df_s_curve <- gen_edges(tri_object = tr1_object_s_curve)
## Compute 2D distances
distance_s_curve <- cal_2d_dist(tr_coord_df = tr_from_to_df_s_curve, 
                             start_x = "x_from", start_y = "y_from", 
                             end_x = "x_to", end_y = "y_to", 
                             select_vars = c("from", "to", "distance"))
# 
# distance_plot
## To find the benchmark value
benchmark_s_curve <- find_lg_benchmark(distance_edges = distance_s_curve, distance_col = "distance")
#distance_plot <- plot_dist(distance_s_curve) +
#   geom_hline(yintercept = benchmark_s_curve, linetype="solid",
#                color = "red", size=0.8, alpha = 0.5) +
#  ylab(expression(d^{(2)})) +
#  theme(axis.text = element_text(size = 5),
#        axis.title = element_text(size = 12))
# ggplot() +
# geom_trimesh(data = df_bin_centroids_pbmc, mapping = aes(x = c_x, y = c_y))
trimesh_s_curve_umap <- vis_lg_mesh(distance_edges = distance_s_curve, benchmark_value = benchmark_s_curve,
tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
trimesh_s_curve_umap <- trimesh_s_curve_umap +
  theme_linedraw() +
  theme(legend.position = "bottom", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) +
  labs(colour = "")
trimesh_removed_s_curve_umap <- vis_rmlg_mesh(distance_edges = distance_s_curve, benchmark_value =
benchmark_s_curve, tr_coord_df = tr_from_to_df_s_curve, distance_col = "distance")
trimesh_removed_s_curve_umap <- trimesh_removed_s_curve_umap +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

df_all_s_curve <- dplyr::bind_cols(s_curve_noise_training |> dplyr::select(-ID), umap_data_with_hb_id)

show_langevitour(df_all_s_curve, df_bin_s_curve, df_bin_centroids_s_curve, benchmark_s_curve, distance_s_curve, "distance", col_start = "x")
```

### 2024-03-31 paper-quollr.tex
Add about main functions

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-2-1.pdf}

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-5-1.pdf}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-2.pdf}

@@

>>>>

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-1-1.pdf}

@@

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-4-1.pdf}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-6-1.pdf}

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-1.pdf} \includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-8-2.pdf}
